Xor-Some Preregistered Study: Analyses

### 2022-05-10 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
plotting and explorations for paper and presentation

d_critical_noDuplicates_zScored_wide <- d_critical_noDuplicates_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  ) 
```

>>>>

d_critical_noDuplicates_zScored_wide <- d_critical_noDuplicates_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  ) 
```
```{r}
# look at raw ratings
# df to be used before applying z-scoring in cell above 
d_critical_zScored_noDuplicates %>%
  filter(block_extended != "target") %>%
  ungroup() %>%
  group_by(block, prior_class, main_type) %>%
  summarize(mean_rating = mean(response)) -> by_type_factor_means

d_critical_zScored_noDuplicates %>%
  filter(block_extended!="target") %>%
  ggplot(., aes(x=main_type, y = response)) +
  geom_jitter(aes(color=prior_class)) +
  geom_point(data = by_type_factor_means, aes(x=main_type, y = mean_rating), color="red") +
  facet_grid(prior_class~block)
```

### 2022-05-10 analysis/main/02_xor-some_prereg_plots_final.Rmd
additional exploratory plotting
plotting and explorations for paper and presentation

Below, some simulations are conducted to create plots for the paper.

>>>>

Below, some simulations are conducted to create plots for the paper. Simulate data similarly to the power analysis, assuming an idealized effect size of 0.2 (somewhat larger than in power analysis, for better visibility). Modify the sampling a little to create low and high condition data.
```{r}
# simulate data from 200 participants with the function from the power analysis
create_fake_data <- function(N = 100, seed = 1) {
  set.seed(seed)
  # optionally changed to 0.25
  beta_rel  <- 0.25
  beta_comp <- 0.25
  beta_pri  <- -0.25
  map_df(
    1:N, 
    function(i) {
      tibble(
        subj    = rep(str_c("subj_",i), times = 8),
        trigger = rep(c('or', 'some'), times= 4), 
        prior = rep(c("low", "high"), each=4),
        rel     = c(rnorm(4, mean=-1), rnorm(4, mean =1)),
       # rel_high = rnorm(4, mean = 0.15),
        comp    = c(rnorm(4, mean=-1), rnorm(4, mean=1)),
       # comp_high=rnorm(4, mean=0.15),
        pri     = c(rnorm(4, mean=1), rnorm(4, mean=-1))
    #    pri_high= rnorm(4, mean=0.15)
      ) %>% 
        group_by(subj) %>% 
        mutate(
          rel     = (rel-mean(rel)) / sd(rel),
          comp    = (comp-mean(comp)) / sd(comp),
          pri     = (pri-mean(pri)) / sd(pri)
        ) %>% 
        ungroup() %>% 
        mutate(
          target  = rnorm(8, beta_rel * rel + beta_comp * comp + beta_pri * pri)
        )
    }
  )
}

fake_data_scatter <- create_fake_data(200, 1234)
```

```{r}
fake_data_long <- fake_data_scatter %>% pivot_longer(
  c(comp, rel, pri), 
               names_to = "Predictors", 
               values_to = "Predictor_response"
) %>%
  mutate(
    prior_categorization = ifelse(Predictor_response <= 0, 0, 1),
    prior_categorization = as.factor(prior_categorization)
  ) %>%
    mutate(
      `Prior categorization` = factor(prior_categorization, levels = c(0,1),
                                         labels=c("low", "high")),
         Trigger = factor(trigger, levels = c("some", "or"), 
                            labels = c("Some", "Or")),
         Predictors = factor(Predictors, levels = c("rel", "comp", "pri"),
                             labels = c("Relevance", "Competence", "Prior"))
    )

# write out for fitting the idealized model in the analysis script
# fake_data_long %>% write_csv("../../data/main/idealized_simulation_long_N200_B025.csv")

fake_data_long %>%
  ggplot(., aes(x = Predictor_response, y = target)) +
  geom_point(aes(color = prior), size = 2, alpha = 0.8) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_grid(cols=vars(Predictors), rows=vars(Trigger)) +
  theme_bw() +
  labs(x = "Participants' factor ratings", y = "Participants' inference likelihood ratings", color = "Prior categorization of vignettes") +
  theme(legend.position = "top",
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 13),
        axis.title = element_text(size = 13),
        strip.text.x = element_text(size = 13),
        strip.text.y = element_text(size = 13)) 
#ggsave("../../writing/paper/images/idealized_data_N200_b025.pdf",  width = 9, height = 7)
```

### 2022-05-10 analysis/main/03_xor-some_prereg_analysis_final.Rmd
plotting and explorations for paper and presentation

model_SI_posteriors_long %>%
  mutate(effect = factor(effect, 
                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
  ggplot(.) + # 
  stat_halfeye(aes(y = effect, x = value, 
                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
                   ), 
               alpha = .7) +
  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
  labs(x = "Effect size", y = "Effect") +
  theme(legend.position = "top",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title = element_text(size = 13),
        strip.text.x = element_text(size = 13),
        strip.text.y = element_text(size = 13))

#ggsave("../../writing/ELM2/posterior-effects.pdf", width=7, height=5)
#ggsave("../../writing/ELM2/posterior-effects-noDupl.pdf", width=7, height=5)
```

@@

model_fake_posteriors_long %>%
  mutate(effect = factor(effect, 
                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
  ggplot(.) + # 
  stat_halfeye(aes(y = effect, x = value, 
                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
                   ), 
               alpha = .7) +
  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
  scale_fill_manual(name = "Direction of effect", values = c("coral", "mediumseagreen")) + # , "yellow3"
  labs(x = "Effect size", y = "Effect") +
  theme(legend.position = "top",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title = element_text(size = 13),
        strip.text.x = element_text(size = 13),
        strip.text.y = element_text(size = 13))

#ggsave("../../writing/paper/images/posterior-effects-idealized-sim.pdf", width=7, height=5)
```

>>>>

model_SI_posteriors_long %>%
  mutate(effect = factor(effect, 
                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
  ggplot(.) + # 
  stat_halfeye(aes(y = effect, x = value, 
                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
                   ), 
               alpha = .7) +
  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) +
  scale_x_continuous(breaks=c(-0.25, 0, 0.25)) +
  labs(x = "Effect size", y = "Effect") +
  theme(legend.position = "top",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title = element_text(size = 13),
        strip.text.x = element_text(size = 13),
        strip.text.y = element_text(size = 13))

#ggsave("../../writing/ELM2/posterior-effects.pdf", width=7, height=5)
#ggsave("../../writing/ELM2/posterior-effects-noDupl.pdf", width=7, height=5)
#ggsave("../../writing/paper/images/posterior-effects-noDupl-breaks.png", width=7, height=5)

@@

model_fake_posteriors_long %>%
  mutate(effect = factor(effect, 
                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
  ggplot(.) + # 
  stat_halfeye(aes(y = effect, x = value, 
                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
                   ), 
               alpha = .7) +
  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
  scale_fill_manual(name = "Direction of effect", values = c("coral", "yellow3", "mediumseagreen")) + 
  scale_x_continuous( breaks=c(-0.25, 0, 0.25)) +
  labs(x = "Effect size", y = "Effect") +
  theme(legend.position = "top",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title = element_text(size = 13),
        strip.text.x = element_text(size = 13),
        strip.text.y = element_text(size = 13))

model_fake_posteriors %>% 
  select(prior_xor, prior_some, 
         competence_xor, competence_some,
         relevance_xor, relevance_some) %>%
  gather(key, val) %>%
  group_by(key) %>% mutate(positive = mean(val > 0.05),
                           negative = mean(val < -0.05),
                           no = mean(val %>% between(-0.05, 0.05))) %>%
  summarise(positive_eff = mean(positive),
            negative_eff = mean(negative),
            no_eff = mean(no))

#ggsave("../../writing/paper/images/posterior-effects-idealized-sim-015.pdf", width=7, height=5)
```
